Sixty-five years of deforestation and forest degradation in Madagascar

### 2017-02-02 deforestmap.Rmd
Writing methodology

We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree percent cover map by @Hansen2013. On Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds). To remove these clouds, we used the 2000 cloud-free tree percent cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover to decide whether to replace cloud pixels with forest or non-forest pixels [@Achard2014]. We thus obtained a cloud-free year 2000 forest map for Madagascar. From this map, using tree cover loss data from 2000 to 2014 by @Hansen2013, we obtained cloud-free forest maps in 2010 and 2014 at 30~m resolution. Finally, we corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990. Doing so, all clouds were removed and we obtained a cloud-free 1990 forest map for Madagascar. These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps.

### 2.2 Deforestation intensity

@@

>>>>

We derived 30m cloud-free forest maps for Madagascar for the years 1990, 2000, 2010 and 2014 combining the 1990 and 2000 forest maps by @Harper2007 and the 2000-2014 tree percent cover map by @Hansen2013. On Harper's map, ~200,000 ha clouds are present over the ~4.2 million ha moist forest ecoregion (4.8% of clouds). To remove these clouds, we used the 2000 cloud-free tree percent cover map provided by @Hansen2013 (also at 30~m resolution) and we chose a threshold of 75% of tree cover to decide whether to replace cloud pixels with forest or non-forest pixels [@Achard2014]. We thus obtained a cloud-free year 2000 forest map for Madagascar. From this map, using tree cover loss data from 2000 to 2014 by @Hansen2013, we obtained cloud-free forest maps in 2010 and 2014 at 30~m resolution. Finally, we corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990. Doing so, all clouds were removed and we obtained a cloud-free 1990 forest map for Madagascar. These maps are freely available at http:/bioscenemada.cirad.fr/forestmaps. On the basis of the forest maps, we computed the mean annual deforestation intensity (in hectares of forest and in percentage) for the five period of time from 1990 to 2014.

### 2.2 Validation

We estimated the accuracy of our approach combining Harper's and Hansen's products to derive cloud-free forest maps after the year 2000 for Madagascar. To do so, we used a deforestation map produced on the period 2005-2013 by a supervised classification of Landsat images at multiple dates using the Random Forest algorithm [@Grinand2013, @Rakotomalala2015]. Following our methodology, we produced a deforestation map on the same period of time (2005-2013) and compute several accuracy indices (overall accuracy, Cohen's Kappa and True Skill Statistics) on the basis of a confusion matrix computed from the two deforestation maps at dates 2005 and 2013.

@@

We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014. We used the approach developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel. We used a moving window of 3x3 pixels (0.81 km<sup2</sup>). Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: interior, perforated, edge, transitional, patch, and undetermined. The interior category can be interpreted as intact forest (see @Potapov2017). We reported the area of forest falling in each category for the 6 years and analysed the evolution of the fragmentation.
